Document S1

### 2025-09-02 latex/docS1.tex
Update the Jackson et al numbers

\subsection*{Analysis of Pango lineage “origination” events}
Notebook

@@

\subsection*{Analysis of the Jackson et al. (2022) recombinants}
Jackson et al. (2022) conducted detailed analyses on some early recombinants involving B.1.1.7 (Alpha), 
including the first designated Pango X lineage XA. 
The authors first performed a targeted search for candidate recombinants 
by scanning sampled sequences for sequence motifs associated with Alpha, 
followed by an analysis to reduce the list of plausible recombinants displaying signs of onward transmission. 
Here we compared our sc2ts detection results of these recombinants with the detection results reported by Jackson et al. (2022).

Jackson et al. (2022) described four groups of recombinants (named A to D) and four singleton recombinants, 
which have distinct inheritance structures (i.e. distinct parental Pango lineages and/or distinct recombination breakpoint intervals). 
All the four groups and 3 out of the 4 singletons were detected as recombinants in the ARG. 
The other singleton was detected as a recombinant by sc2ts, 
but it was not added to the ARG due to a high HMM cost (two breakpoints with X mutations). 
The sc2ts detection results are overall concordant with the results of Jackson et al. (2022) in terms of 
parental Pango lineages (agreeing to at least three-level resolution) and breakpoint intervals (Table S1).

Table S1: Sc2ts detection results for the recombinants documented by Jackson et al. (2022).
https://github.com/jeromekelleher/sc2ts-paper/issues/293

>>>>

\subsection*{Analysis of Pango lineage ``origination'' events}
Notebook

@@

\subsection*{Analysis of the Jackson et al. (2021) recombinants}
Jackson et al. (2021) conducted detailed analyses on some early recombinants
involving B.1.1.7 (Alpha), including the first designated Pango X lineage, XA.
The authors first performed a targeted search for candidate recombinants by
scanning sampled sequences for sequence motifs associated with Alpha, followed
by an analysis to reduce the list of plausible recombinants displaying signs of
onward transmission. Here we compared our sc2ts detection results of these
recombinants with the detection results reported by Jackson et al.

Jackson et al.\ described four groups of recombinants (named A to D) and
four singleton recombinants.
Of the 16 samples analysed by Jackson et al, 12 were included in the ARG;
the remaining 4 were excluded because the sequences did not pass the 
QC filtering criteria, or due to a high HMM cost. Groups A-D are summarised 
in Table~\ref{tab:jackson} and show excellent concordance in terms 
of the parental Pango lineages and breakpoint intervals. All four groups
are directly associated with strongly supported recombination events 
in the ARG in terms of the number of averted mutations 
(A: 16, B: 6, C: 30, D: 18). 
The singleton CAMC-CBA018 has parents
B.1.177+B.1.1.7 and breakpoint interval 16,177--20,133 
in the ARG 
and is strongly supported by 14 averted mutations.
The Jackson et al.\ results are almost identical with the same parent 
lineages, and breakpoint interval of interval 20,390--21,256.
The singleton MILK-103C712 
has parents B.1.177.16+B.1.1.7 and a breakpoint interval of 
24915--27972 
in the ARG, again strongly supported by 11 averted mutations.
The Jackson et al.\ results for this lineage 
has parents B.1.177.17+B.1.1.7 but two breakpoints
409--446 and 26,802--27,878.

\begin{table} \centering
\caption{\label{tab:jackson}Comparison of recombination breakpoint
intervals and parent lineages for Groups A-D
reported by Jackson et al.\ with the corresponding
recombination events in the sc2ts ARG.
The second column gives the number of sequences in the group.
See the text for details of the groups and the sequences included.
limited to the samples considered by Jackson et al.
The 3SEQ coordinates
reported by Jackson et al.\ have been altered as follows: we
add one to both left and right coordinates to correspond to the
\texttt{tskit} definition of inheritance on either side of a breakpoint,
and add one to the right coordinate to make the intervals
right-exclusive. }
\begin{tabular}{lllr@{--}lr@{+}l}
\toprule
Group        & Sequences & Method & \multicolumn{2}{c}{Interval}
    & \multicolumn{2}{r}{Parent lineages} \\
\midrule
A (XA)       & 2   & Jackson        &  21,256&21,615 & B.1.177&B.1.1.7 \\
             &     &\texttt{sc2ts} &  20,411&21,765 & B.1.177.18&B.1.1.7 \\
\cmidrule{3-7}
B            & 2   & Jackson        &  6,529&6,955 & B.1.36.28&B.1.1.7  \\
             &     &\texttt{sc2ts} &   6,529&6,954 & B.1.36.28&B.1.1.7  \\
\cmidrule{3-7}
C            & 3   &Jackson         &  25,997&27,443 &  B.1.1.7&B.1.221 \\
             &     & \texttt{sc2ts} &  25,997&27,972 &  B.1.1.7&B.1.221.1 \\
\cmidrule{3-7}
D            & 3   & Jackson        &  21,576&23,064 &  B.1.36.17&B.1.1.7 \\
             &     & \texttt{sc2ts} &  22,445&23,063 &  B.1.36.39&B.1.1.7 \\
\bottomrule
\end{tabular}
\end{table}
